---
title: "exercise-05"
format: html
---

# Challenge 1

## Step 1

``` {r}
library(tidyverse)
d <- tibble(read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/IMDB-movies.csv"))
```

## Step 2

``` {r}
d <- d %>%
  filter(startYear >= 1920 & startYear <= 1979 & runtimeMinutes >= 60 & runtimeMinutes <= 180) %>%
  mutate(decade = paste(floor((startYear/10)) * 10 - 1900, "s", sep = ""))
```

## Step 3

``` {r}
p <- ggplot(data = d, aes(x = runtimeMinutes)) +
  geom_histogram() +
  facet_wrap(facets = vars(decade))
p
```

## Step 4

``` {r}
results <- d %>%
  group_by(decade) %>%
  summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
results
```

## Step 5

``` {r}
samp <- d %>%
  group_by(decade) %>%
  slice_sample(n = 100)
samp_info <- samp %>%
  summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
samp_info
```

## Step 6

``` {r}
samp_info <- samp_info %>%
  mutate(standardError = standardDeviation / sqrt(100))
samp_info
```

## Step 8

``` {r}
samp_dist <- tibble(decade = character(), mean = numeric(), standardDeviation = numeric())
for(i in 1:1000) {
  samp_1000 <- d %>%
    group_by(decade) %>%
    slice_sample(n = 100)
  samp_1000_info <- samp_1000 %>%
    group_by(decade) %>%
    summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
  samp_dist <- samp_dist %>%
    add_row(samp_1000_info)
}
samp_dist <- samp_dist %>%
  remove_missing()
```

## Step 9

``` {r}
samp_dist_info <- samp_dist %>%
  group_by(decade) %>%
  summarise(Mean = mean(mean), standardDeviation = sd(mean))
samp_p <- ggplot(data = samp_dist, aes(x = mean)) +
  geom_histogram() +
  facet_wrap(facets = vars(decade))
samp_p
```

# Challenge 2

## Step 1

``` {r}
z <- tibble(read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/zombies.csv"))
```

## Step 2

``` {r}
pop_sd <- function(pop) {
  sd <- sd(pop) * sqrt((length(pop) - 1)/length(pop))
  return(sd)
}
z_long <- z %>%
  pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value")
z_info <- z_long %>%
  group_by(variable) %>%
  summarise(mean = mean(value), standardDeviation = pop_sd(value))
```

## Step 3

``` {r}
z_height <- ggplot(data = z, aes(x = gender, y = height)) +
  geom_boxplot()
z_height
```

``` {r}
z_weight <- ggplot(data = z, aes(x = gender, y = weight)) +
  geom_boxplot()
z_weight
```

``` {r}
z_age <- ggplot(data = z, aes(x = gender, y = age)) +
  geom_boxplot()
z_age
```

``` {r}
z_zombies_killed <- ggplot(data = z, aes(x = gender, y = zombies_killed)) +
  geom_boxplot()
z_zombies_killed
```

``` {r}
z_years_of_education <- ggplot(data = z, aes(x = gender, y = years_of_education)) +
  geom_boxplot()
z_years_of_education
```

## Step 4

``` {r}
z_female <- z %>%
  filter(gender == "Female")
z_male <- z %>%
  filter(gender == "Male")
z_age_height <- ggplot(data = z_female, aes(x = age, y = height)) +
  geom_point(colour = "pink") +
  geom_point(data = z_male, colour = "blue")
z_age_height
```

``` {r}
z_age_weight <- ggplot(data = z_female, aes(x = age, y = weight)) +
  geom_point(colour = "pink") +
  geom_point(data = z_male, colour = "blue")
z_age_weight
```

Height and weight seem to follow similar trends--they're both higher in males, and they increase with age.

## Step 5

``` {r}
z_height_hist <- ggplot(data = z, aes(x = height)) +
  geom_histogram()
z_height_hist
```

``` {r}
z_weight_hist <- ggplot(data = z, aes(x = weight)) +
  geom_histogram()
z_weight_hist
```

``` {r}
z_age_hist <- ggplot(data = z, aes(x = age)) +
  geom_histogram()
z_age_hist
```

``` {r}
z_zombies_killed_hist <- ggplot(data = z, aes(x = zombies_killed)) +
  geom_histogram(bins = 10)
z_zombies_killed_hist
```

``` {r}
z_years_of_education_hist <- ggplot(data = z, aes(x = years_of_education)) +
  geom_histogram(bins = 9)
z_years_of_education_hist
```

``` {r}
z_height_qq <- ggplot(data = z, aes(sample = height)) +
  geom_qq()
z_height_qq
```

``` {r}
z_weight_qq <- ggplot(data = z, aes(sample = weight)) +
  geom_qq()
z_weight_qq
```

``` {r}
z_age_qq <- ggplot(data = z, aes(sample = age)) +
  geom_qq()
z_age_qq
```

``` {r}
z_zombies_killed_qq <- ggplot(data = z, aes(sample = zombies_killed)) +
  geom_qq()
z_zombies_killed_qq
```

``` {r}
z_years_of_education_qq <- ggplot(data = z, aes(sample = years_of_education)) +
  geom_qq()
z_years_of_education_qq
```

Height, weight, and age seem to be drawn from a normal distribution, while number of zombies killed and years of education do not.

## Step 6

``` {r}
z_sample <- z %>%
  slice_sample(n = 50)
z_sample_long <- z_sample %>%
  pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value")
z_sample_info <- z_sample_long %>%
  group_by(variable) %>%
  summarise(mean = mean(value), standardDeviation = sd(value), standardError = sd(value)/sqrt(length(value)), CI95Lower = mean + qnorm(0.025) * standardError, CI95Upper = mean + qnorm(0.975) * standardError)
z_sample_info
```

## Step 7

``` {r}
z_dist <- z_sample_info
for(i in 1:199) {
  z_200 <- z %>%
    slice_sample(n = 50)
  z_200_long <- z_200 %>%
    pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value")
  z_200_info <- z_200_long %>%
    group_by(variable) %>%
    summarise(mean = mean(value), standardDeviation = sd(value), standardError = sd(value)/sqrt(length(value)), CI95Lower = mean + qnorm(0.025) * standardError, CI95Upper = mean + qnorm(0.975) * standardError)
  z_dist <- z_dist %>%
    add_row(z_200_info)
}
z_dist_info <- z_dist %>%
  group_by(variable) %>%
  summarise(Mean = mean(mean), standardDeviation = sd(mean))
```

The means of the sampling distribution are similar to the means of the first sample taken, and the sampling distribution's standard deviation is also comparable to the first sample's standard error.

## Step 8

``` {r}
z_dist_height <- z_dist %>%
  filter(variable == "height")
z_dist_height_hist <- ggplot(data = z_dist_height, aes(mean)) +
  geom_histogram()
z_dist_height_hist
```

``` {r}
z_dist_weight <- z_dist %>%
  filter(variable == "weight")
z_dist_weight_hist <- ggplot(data = z_dist_weight, aes(mean)) +
  geom_histogram()
z_dist_weight_hist
```

``` {r}
z_dist_age <- z_dist %>%
  filter(variable == "age")
z_dist_age_hist <- ggplot(data = z_dist_age, aes(mean)) +
  geom_histogram()
z_dist_age_hist
```

``` {r}
z_dist_zombies_killed <- z_dist %>%
  filter(variable == "zombies_killed")
z_dist_zombies_killed_hist <- ggplot(data = z_dist_zombies_killed, aes(mean)) +
  geom_histogram()
z_dist_zombies_killed_hist
```

``` {r}
z_dist_years_of_education <- z_dist %>%
  filter(variable == "years_of_education")
z_dist_years_of_education_hist <- ggplot(data = z_dist_years_of_education, aes(mean)) +
  geom_histogram()
z_dist_years_of_education_hist
```

The variables all seem more normally distributed as sampling distributions.

## Step 9

``` {r}
z_dist_realCI <- z_dist %>%
  group_by(variable) %>%
  summarise(realCI95Lower = quantile(mean, probs = 0.025), realCI95Upper = quantile(mean, probs = 0.975))
z_dist_real_CI
```

The 95% confidence intervals seem pretty similar between the sampling distribution and the individual samples.

## Step 10

``` {r}
bootstrap <- function(data, n_boot = 1000) {
  boot <- vector(length = n_boot)
  n <- length(data)
  for(i in 1:n_boot) {
    boot[[i]] <- mean(sample(data, n, replace = TRUE))
    return(boot)
  }
}
z_sample_height <- z_sample_long %>%
  filter("variable" == "height")
z_sample_weight <- z_sample_long %>%
  filter("variable" == "weight")
z_sample_age <- z_sample_long %>%
  filter("variable" == "height")
z_sample_zombies_killed <- z_sample_long %>%
  filter("variable" == "zombies_killed")
z_sample_years_of_education <- z_sample_long %>%
  filter("variable" == "years_of_education")

boot_height <- bootstrap(z_sample_height$value)
print(z_sample_height$value)
z_boot <- tibble("variable" = "height", "CI95Lower" = quantile(boot_height, probs = 0.025), "CI95Upper" = quantile(boot_height, probs = 0.975))

```