---
title: "exercise-05"
format: html
---

# Challenge 1

## Step 1

``` {r}
library(tidyverse)
library(mosaic)
d <- tibble(read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/IMDB-movies.csv"))
```

## Step 2

``` {r}
d <- d %>%
  filter(startYear >= 1920 & startYear <= 1979 & runtimeMinutes >= 60 & runtimeMinutes <= 180) %>%
  mutate(decade = paste(floor((startYear/10)) * 10 - 1900, "s", sep = "")) # filters the dataset to only include movies with 1-3 hour runtimes and identifies which decade they were released in
```

## Step 3

``` {r}
p <- ggplot(data = d, aes(x = runtimeMinutes)) +
  geom_histogram() +
  facet_wrap(facets = vars(decade))
p # histogram of movies between 1 and 3 hours separated by decade
```

## Step 4

``` {r}
results <- d %>%
  group_by(decade) %>%
  summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
results # summarizes each decade's runtimes
```

## Step 5

``` {r}
samp <- d %>%
  group_by(decade) %>%
  slice_sample(n = 100)
samp_info <- samp %>%
  summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
samp_info # summarizes runtimes of 100-movie samples from each decade
```

## Step 6

``` {r}
samp_info <- samp_info %>%
  mutate(standardError = standardDeviation / sqrt(100))
samp_info # added standard error to above table
```

## Step 8

``` {r}
samp_dist <- tibble(decade = character(), mean = numeric(), standardDeviation = numeric())
for(i in 1:1000) {
  samp_1000 <- d %>%
    group_by(decade) %>%
    slice_sample(n = 100)
  samp_1000_info <- samp_1000 %>%
    group_by(decade) %>%
    summarise(mean = mean(runtimeMinutes), standardDeviation = sd(runtimeMinutes))
  samp_dist <- samp_dist %>%
    add_row(samp_1000_info)
}
samp_dist <- samp_dist %>%
  remove_missing() # logs all samples for the 1000-sample distribution of 100 movies each, for each decade
```

## Step 9

``` {r}
samp_dist_info <- samp_dist %>%
  group_by(decade) %>%
  summarise(Mean = mean(mean), standardDeviation = sd(mean))
samp_p <- ggplot(data = samp_dist, aes(x = mean)) +
  geom_histogram() +
  facet_wrap(facets = vars(decade))
samp_p # plots histograms of the above distributions
```

## Step 10

The standard error for runtime seems pretty similar compared to the standard error for the first sample for the distribution, the population standard deviation, and the standard deviation for the entire sampling distribution.

# Challenge 2

## Step 1

``` {r}
z <- tibble(read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/zombies.csv"))
```

## Step 2

``` {r}
pop_sd <- function(pop) {
  sd <- sd(pop) * sqrt((length(pop) - 1)/length(pop))
  return(sd)
} # function for converting the native sample standard deviation function to a population standard deviation
z_long <- z %>%
  pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value") # turns each variable we want to look at into one column, which can be filtered by what variable that specific row has in its "value" column using the also-new "variable" column
z_info <- z_long %>%
  group_by(variable) %>%
  summarise(mean = mean(value), standardDeviation = pop_sd(value)) # generates a table that gives a brief summary of the entire table's/population's different quantized attributes
```

## Step 3

``` {r}
z_height <- ggplot(data = z, aes(x = gender, y = height)) +
  geom_boxplot()
z_height # boxplot for height, separated by gender
```

``` {r}
z_weight <- ggplot(data = z, aes(x = gender, y = weight)) +
  geom_boxplot()
z_weight
```

``` {r}
z_age <- ggplot(data = z, aes(x = gender, y = age)) +
  geom_boxplot()
z_age
```

``` {r}
z_zombies_killed <- ggplot(data = z, aes(x = gender, y = zombies_killed)) +
  geom_boxplot()
z_zombies_killed
```

``` {r}
z_years_of_education <- ggplot(data = z, aes(x = gender, y = years_of_education)) +
  geom_boxplot()
z_years_of_education
```

## Step 4

``` {r}
z_female <- z %>%
  filter(gender == "Female")
z_male <- z %>%
  filter(gender == "Male")
z_age_height <- ggplot(data = z_female, aes(x = age, y = height)) +
  geom_point(colour = "pink") +
  geom_point(data = z_male, colour = "blue")
z_age_height # scatterplot of height; the different colors differentiate gender for each subject
```

``` {r}
z_age_weight <- ggplot(data = z_female, aes(x = age, y = weight)) +
  geom_point(colour = "pink") +
  geom_point(data = z_male, colour = "blue")
z_age_weight
```

Height and weight seem to follow similar trends--they're both higher in males, and they increase with age.

## Step 5

``` {r}
z_height_hist <- ggplot(data = z, aes(x = height)) +
  geom_histogram()
z_height_hist # histogram of the entire population's heights
```

``` {r}
z_weight_hist <- ggplot(data = z, aes(x = weight)) +
  geom_histogram()
z_weight_hist
```

``` {r}
z_age_hist <- ggplot(data = z, aes(x = age)) +
  geom_histogram()
z_age_hist
```

``` {r}
z_zombies_killed_hist <- ggplot(data = z, aes(x = zombies_killed)) +
  geom_histogram(bins = 10)
z_zombies_killed_hist
```

``` {r}
z_years_of_education_hist <- ggplot(data = z, aes(x = years_of_education)) +
  geom_histogram(bins = 9)
z_years_of_education_hist
```

``` {r}
z_height_qq <- ggplot(data = z, aes(sample = height)) +
  geom_qq()
z_height_qq # quantile-quantile plot of the entire population's heights. The X-axis is what the expected value for a given subject percentile's height would be, assuming the data is normally distributed; the Y-axis is the actual value for that subject percentile's height. The more linear the graph is, the more normal the distribution.
```

``` {r}
z_weight_qq <- ggplot(data = z, aes(sample = weight)) +
  geom_qq()
z_weight_qq
```

``` {r}
z_age_qq <- ggplot(data = z, aes(sample = age)) +
  geom_qq()
z_age_qq
```

``` {r}
z_zombies_killed_qq <- ggplot(data = z, aes(sample = zombies_killed)) +
  geom_qq()
z_zombies_killed_qq
```

``` {r}
z_years_of_education_qq <- ggplot(data = z, aes(sample = years_of_education)) +
  geom_qq()
z_years_of_education_qq
```

Height, weight, and age seem to be drawn from a normal distribution, while number of zombies killed and years of education do not.

## Step 6

``` {r}
z_sample <- z %>%
  slice_sample(n = 50)
z_sample_long <- z_sample %>%
  pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value")
z_sample_info <- z_sample_long %>%
  group_by(variable) %>%
  summarise(mean = mean(value), standardDeviation = sd(value), standardError = sd(value)/sqrt(length(value)), CI95Lower = mean + qnorm(0.025) * standardError, CI95Upper = mean + qnorm(0.975) * standardError)
z_sample_info # summarizes the variables of a 50-subject sample
```

## Step 7

``` {r}
z_dist <- z_sample_info
for(i in 1:199) {
  z_200 <- z %>%
    slice_sample(n = 50)
  z_200_long <- z_200 %>%
    pivot_longer(cols = c("height", "weight", "age", "zombies_killed", "years_of_education"), names_to = "variable", values_to = "value")
  z_200_info <- z_200_long %>%
    group_by(variable) %>%
    summarise(mean = mean(value), standardDeviation = sd(value), standardError = sd(value)/sqrt(length(value)), CI95Lower = mean + qnorm(0.025) * standardError, CI95Upper = mean + qnorm(0.975) * standardError)
  z_dist <- z_dist %>%
    add_row(z_200_info)
} # table containing all the info from the sampling distribution
z_dist_info <- z_dist %>%
  group_by(variable) %>%
  summarise(Mean = mean(mean), standardDeviation = sd(mean)) # summary of the sampling distribution's variables
```

The means of the sampling distribution are similar to the means of the first sample taken, and the sampling distribution's standard deviation is also comparable to the first sample's standard error.

## Step 8

``` {r}
z_dist_height <- z_dist %>%
  filter(variable == "height")
z_dist_height_hist <- ggplot(data = z_dist_height, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(z_dist_height$mean, probs = 0.025)) +
  geom_vline(xintercept = quantile(z_dist_height$mean, probs = 0.975))
z_dist_height_hist # histogram of the sampling distribution's sample means for height with empirically determined 95% confidence intervals included
```

``` {r}
z_dist_weight <- z_dist %>%
  filter(variable == "weight")
z_dist_weight_hist <- ggplot(data = z_dist_weight, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(z_dist_weight$mean, probs = 0.025)) +
  geom_vline(xintercept = quantile(z_dist_weight$mean, probs = 0.975))
z_dist_weight_hist
```

``` {r}
z_dist_age <- z_dist %>%
  filter(variable == "age")
z_dist_age_hist <- ggplot(data = z_dist_age, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(z_dist_age$mean, probs = 0.025)) +
  geom_vline(xintercept = quantile(z_dist_age$mean, probs = 0.975))
z_dist_age_hist
```

``` {r}
z_dist_zombies_killed <- z_dist %>%
  filter(variable == "zombies_killed")
z_dist_zombies_killed_hist <- ggplot(data = z_dist_zombies_killed, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(z_dist_zombies_killed$mean, probs = 0.025)) +
  geom_vline(xintercept = quantile(z_dist_zombies_killed$mean, probs = 0.975))
z_dist_zombies_killed_hist
```

``` {r}
z_dist_years_of_education <- z_dist %>%
  filter(variable == "years_of_education")
z_dist_years_of_education_hist <- ggplot(data = z_dist_years_of_education, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(z_dist_years_of_education$mean, probs = 0.025)) +
  geom_vline(xintercept = quantile(z_dist_years_of_education$mean, probs = 0.975))
z_dist_years_of_education_hist
```

The variables all seem more normally distributed as sampling distributions.

## Step 9

``` {r}
z_dist_realCI <- z_dist %>%
  group_by(variable) %>%
  summarise(Mean = mean(mean), standardDeviation = sd(mean), realCI95Lower = quantile(mean, probs = 0.025), realCI95Upper = quantile(mean, probs = 0.975))
z_dist_realCI # table containing empirically determined values for each variable's 95% confidence intervals
```

The 95% confidence intervals seem pretty similar between the sampling distribution and the individual samples.

## Step 10

``` {r}
n_boot <- 1000
boot_height <- do(n_boot) * mean(~sample(z_sample$height, size = length(z_sample$height), replace = TRUE)) # bootstraps 1000 samples of height taken from our first sample
lower_boot_height <- quantile(boot_height$mean, probs = 0.025)
upper_boot_height <- quantile(boot_height$mean, probs = 0.975)
boot_height_plot <- ggplot(data = boot_height, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = lower_boot_height) +
  geom_vline(xintercept = upper_boot_height)
boot_height_plot # histogram of the bootstrap sample means with empirically determined 95% confidence intervals
```

``` {r}
boot_weight <- do(n_boot) * mean(~sample(z_sample$weight, size = length(z_sample$weight), replace = TRUE))
lower_boot_weight <- quantile(boot_weight$mean, probs = 0.025)
upper_boot_weight <- quantile(boot_weight$mean, probs = 0.975)
boot_weight_plot <- ggplot(data = boot_weight, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = lower_boot_weight) +
  geom_vline(xintercept = upper_boot_weight)
boot_weight_plot
```

``` {r}
boot_age <- do(n_boot) * mean(~sample(z_sample$age, size = length(z_sample$age), replace = TRUE))
lower_boot_age <- quantile(boot_age$mean, probs = 0.025)
upper_boot_age <- quantile(boot_age$mean, probs = 0.975)
boot_age_plot <- ggplot(data = boot_age, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = lower_boot_age) +
  geom_vline(xintercept = upper_boot_age)
boot_age_plot
```

``` {r}
boot_zombies_killed <- do(n_boot) * mean(~sample(z_sample$zombies_killed, size = length(z_sample$zombies_killed), replace = TRUE))
lower_boot_zombies_killed <- quantile(boot_zombies_killed$mean, probs = 0.025)
upper_boot_zombies_killed <- quantile(boot_zombies_killed$mean, probs = 0.975)
boot_zombies_killed_plot <- ggplot(data = boot_zombies_killed, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = lower_boot_zombies_killed) +
  geom_vline(xintercept = upper_boot_zombies_killed)
boot_zombies_killed_plot
```

``` {r}
boot_years_of_education <- do(n_boot) * mean(~sample(z_sample$years_of_education, size = length(z_sample$years_of_education), replace = TRUE))
lower_boot_years_of_education <- quantile(boot_years_of_education$mean, probs = 0.025)
upper_boot_years_of_education <- quantile(boot_years_of_education$mean, probs = 0.975)
boot_years_of_education_plot <- ggplot(data = boot_years_of_education, aes(mean)) +
  geom_histogram() +
  geom_vline(xintercept = lower_boot_years_of_education) +
  geom_vline(xintercept = upper_boot_years_of_education)
boot_years_of_education_plot
```

``` {r}
z_boot_info <- tibble(
  variable = c("height", "weight", "age", "zombies_killed", "years_of_education"),
  Mean = c(mean(boot_height$mean), mean(boot_weight$mean), mean(boot_age$mean), mean(boot_zombies_killed$mean), mean(boot_years_of_education$mean)),
  standardDeviation = c(sd(boot_height$mean), sd(boot_weight$mean), sd(boot_age$mean), sd(boot_zombies_killed$mean), sd(boot_years_of_education$mean)),
  CI95Lower = c(lower_boot_height, lower_boot_weight, lower_boot_age, lower_boot_zombies_killed, lower_boot_years_of_education),
  CI95Upper = c(upper_boot_height, upper_boot_weight, upper_boot_age, upper_boot_zombies_killed, upper_boot_years_of_education)
)
z_boot_info # table containing info about the bootstrapped sample distribution
```

The confidence intervals generated via bootstrapping are very similar to the CIs generated by sampling distribution in Step 9.